A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2023-07-02 paper/fascd.tex
remove an unneeded footnote

The mesh hierarchy and FE spaces can be constructed by standard nested refinement.\footnote{The dimension-dependent details of mesh refinement are suppressed here.  Section \ref{sec:results} only uses $P_1$ elements for intervals, triangles, and tetrahedra, in dimensions $d=1,2,3$ respectively, within the Firedrake framework \cite{Rathgeberetal2016}.}  Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements with union equal to $\overline{\Omega}$ and maximum edge length (characteristic mesh size) $h_0>0$.  We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.  For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ so that each $\mathcal{T}^{j-1}$ element becomes $2^d$ elements in $\mathcal{T}^j$ with maximum edge length $h_j = 2^{-j} h_0$.  We call $J\ge 0$ the \emph{depth} of the hierarchy, $J+1$ the \emph{number of levels}, and $\mathcal{T}^J$ the \emph{finest-level mesh}.  Let $m_j$ be the number of nodes in $\mathcal{T}^j$, so there are $m_j$ \emph{degrees of freedom} at the $j$th level.  Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
\begin{equation}
\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}=W^{1,p}(\Omega).  \label{eq:fe:nestedspaces}
\end{equation}
On the finest-level mesh let $g_D^J \in \mathcal{V}^J$ satisfy $g_D^J = g_D$ on $\partial_D \Omega$; this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear.

In the VI material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.  Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},
\begin{equation}
\bw \ge \bz \quad \implies \quad w \ge z \label{eq:nodalmonotonicity}
\end{equation}
for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.  However, for $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.  Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign hat functions.

>>>>

The mesh hierarchy and FE spaces can be constructed by standard nested refinement.  Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements with union equal to $\overline{\Omega}$ and maximum edge length (characteristic mesh size) $h_0>0$.  We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.  For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ so that each $\mathcal{T}^{j-1}$ element becomes $2^d$ elements in $\mathcal{T}^j$ with maximum edge length $h_j = 2^{-j} h_0$.  We call $J\ge 0$ the \emph{depth} of the hierarchy, $J+1$ the \emph{number of levels}, and $\mathcal{T}^J$ the \emph{finest-level mesh}.  Let $m_j$ be the number of nodes in $\mathcal{T}^j$, so there are $m_j$ \emph{degrees of freedom} at the $j$th level.  Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
\begin{equation}
\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}=W^{1,p}(\Omega).  \label{eq:fe:nestedspaces}
\end{equation}
On the finest-level mesh let $g_D^J \in \mathcal{V}^J$ satisfy $g_D^J = g_D$ on $\partial_D \Omega$; this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear.

In the material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.  Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},
\begin{equation}
\bw \ge \bz \quad \implies \quad w \ge z \label{eq:nodalmonotonicity}
\end{equation}
for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.  However, for $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.  Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign hat functions.
